\section{Nomenclature and Notation}

\begin{itemize}
\item \textbf{Aether resonance:} substrate-local coupling in pattern space $(S)$ projected as FTL in spacetime $(M)$.
\item \textbf{Pattern space $(S)$:} space of local data structures; distance $(d_\sigma)$ measures isomorphic similarity, \textbf{has length dimension (meters)}.
\item \textbf{$\varepsilon$:} dimensionless coupling strength ($\sim 10^{-15}$).
\item \textbf{$\alpha$:} gravitational coupling factor. \textbf{In main text set $\alpha\equiv1$} to satisfy Bianchi identity at $\pm J_\sigma^\nu$.
\item \textbf{$\hbar\omega_0$:} characteristic energy scale ($\sim 10^{-23}$ J at 10 GHz).
\item \textbf{$\lambda_\sigma$:} coherence length in S, \textbf{length dimension (meters)}.
\item \textbf{$\ell_0$:} embedding scale giving $(d_\sigma)$ length dimension ($\sim 1~\mu$m).
\item \textbf{Q ($\mathcal{Q}$):} coherence/quality, dimensionless, $0 \leq Q \leq 1$ ($\sim 10^{-10}$ standard, $\sim 10^{-2}$ optimized).
\item \textbf{K:} similarity kernel, $(K = \exp[-d_\sigma/\lambda_\sigma])$, dimensionless.
\item \textbf{$\tilde{\mathcal K}$:} dimensionless cost factor $=P_{\rm pump}/\hbar\omega_0$ [$\mathrm{s^{-1}}$].
\item \textbf{$O_S$:} selection operator, mass dimension $\Delta > 4$.
\item \textbf{$J^\nu_\sigma$:} four-current from S-interaction, $[J^\nu_\sigma]$ = energy$\cdot$density / time.
\item \textbf{$T^{\mu\nu}_S$:} energy-momentum tensor for S-sector.
\item \textbf{$\chi(\sigma,T)$:} S-mediator field propagating locally in S and mediating resonance.
\item \textbf{Degeneracy dilution:} weakening ($\sim 1/N$) when many matches compete.
\item \textbf{$\tau$:} foliation scalar (khronon) defining substrate's preferred time.
\item \textbf{$u^\mu$:} unit timelike vector along preferred frame.
\end{itemize}

