\section{Experimental Protocols (No-Loophole) and Parameter Inference}

All protocols are \textbf{pre-registered} (Open Science Framework), timestamped, and data/analysis published openly.

\textbf{Statistical method box:} We pre-register decision rules and adjust for \textbf{multiple comparisons} (e.g., \textbf{Holm–Bonferroni} or FDR correction) over distance levels (6 levels in §7.3), control protocols (match/mismatch, on/off), and experiments (E1–E3) to control family-wise error rate at $\alpha=0.05$. The distance ladder pilot (see §7.3) is published with its own DOI and used as advance calibration.

\subsection{E1. Neuromorphic Ansible (Information)}

\textbf{Setup.}
\begin{itemize}
\item Two photonic/electronic reservoirs (3D RC networks, $N \sim 10^3$ nodes),
\item Training on identical dataset (MNIST / audio / video-snippet),
\item \textbf{Triple-Faraday}, optical isolation (fiber-air-gap, optical isolators), battery power,
\item Independent atomic clocks (GPS-disciplined or crystal-OCXO, jitter $< 1$ ns),
\item \textbf{Spacelike separation} ($> 1$ km, light-travel-time $> 3~\mu$s).
\end{itemize}

\textbf{Protocol.}
\begin{enumerate}
\item \textbf{Pre-commitment:} Publish hash (SHA-256) of codebook + time schedule (\textbf{commit–reveal}).
\item[1a.] \textbf{Blind pilot (distance ladder):} Run a separate, blindly labeled level series $(N=6)$ and freeze $(\lambda_\sigma,\ell_0)$ before main run.
\item \textbf{Distance ladder:} Test 6 levels (§7.3), $10^9$ bits per level.
\item \textbf{Delayed choice:} QRNG selects codebook/times after commit, before transmission.
\item \textbf{Blind reception:} Receiver runs correlation against codebook without sender access.
\item \textbf{Sham blocks:} 20\% of blocks are ``off'' (sender inactive) $\to$ test for false-positive (ghost blocks).
\item \textbf{Cosmic veto:} Muon detector; reject if $(N_\mu > 10 / \text{m}^2 / \text{s})$.
\end{enumerate}

\textbf{Analysis.}
\begin{itemize}
\item Calculate BER per level,
\item Bayes factor ($H_1$: resonance vs. $H_0$: chance + leakage),
\item Sequential Probability Ratio Test (SPRT) for early stop,
\item Permutation tests ($\geq 10^6$) for p-value,
\item \textbf{Holm-Bonferroni correction} for 6 levels.
\end{itemize}

\textbf{Goal:} $\Delta$BER $\sim 10^{-3}$ (match vs. mismatch) with cross-correlation/coherence as primary statistic, $(p < 10^{-6})$ (corrected).

\textbf{Inference:} Null result maps to $(\varepsilon \lambda_\sigma Q)$ via (8.3-8.4).

\subsection{E2. Energy Tunnel (Energy)}

\textbf{Setup.}
\begin{itemize}
\item Two identical superconducting microwave cavities or metamaterial rigs $(Q_{cav} \sim 10^6)$,
\item \textbf{Cryo-calorimetry} $(T \sim 10$ mK, $\delta T \sim 0.1~\mu$K $\to \delta E \sim 10^{-26}$ J),
\item Separation $> 1$ km (light-travel-time $> 3~\mu$s),
\item \textbf{Rotation platform} (0.1 rpm $\to$ scan preferred frame for sidereal modulation).
\end{itemize}

\textbf{Protocol.}
\begin{enumerate}
\item \textbf{Matching test:} Vary internal geometry (0\%, 50\%, 100\% match).
\item \textbf{Pump modulation:} A pumped with $(P_{pump} = 1~\mu$W), period 100 s on/off; B below threshold.
\item \textbf{Phase-locking on/off:} Control coherence $(Q)$.
\item \textbf{Thermostat on/off:} Alternate active vs. passive temperature control.
\item \textbf{Latency scan:} Correlate $(\Delta T_B(t))$ with $(P_A(t-\delta))$, scan $\delta \in [-10~\mu\text{s}, +10~\mu\text{s}]$.
\item \textbf{Grav-sidekick:} Torsion pendulum or cavity in B (sensitivity $\sim 10^{-15}$ m/s$^2$); report expected $(\Delta \Phi_g)$ even if below sensitivity.
\item \textbf{Momentum test:} Verify (3.7) experimentally with precision force meters (no traces of reactionless force).
\item \textbf{Rotation test:} Scan preferred-frame direction for sidereal modulation (11.2).
\end{enumerate}

\textbf{Goals:}
\begin{itemize}
\item $(\Delta E_A + \Delta E_B)$ deviates from zero with $(|\Delta E| > 10^{-25}$ J) (ambitious scenario),
\item Correlation with matching grade $(r > 0.8)$,
\item FTL latency $(\delta < -3~\mu$s) vs. thermal leakage $(\delta > 0)$,
\item Sidereal modulation $(A \geq 10^{-20})$.
\end{itemize}

\textbf{Analysis:}
\begin{itemize}
\item Cross-correlation and regression model with (10) and (11.2),
\item Bayes factors, permutation tests,
\item \textbf{Multiple-test correction} for match/mismatch, on/off, and rotation conditions.
\end{itemize}

\textbf{Null bound:} If $(|\Delta E| < 10^{-26}$ J) after $10^6$ s $\to (\varepsilon \omega_0 Q < 10^{-8}$ Hz). Rotation null $\to (\varepsilon Q < 10^{-20})$.

\subsection{E3. Chaos-to-Chaos}

\textbf{Setup.}
\begin{itemize}
\item Two turbulent flows (Rayleigh-Bénard cells, $L = 10$ cm) or reaction-diffusion systems (identical geometry),
\item Laser Doppler velocimetry or imaging (1 kHz sampling),
\item Drive: modulated heat flux or chemical concentration with variable complexity.
\end{itemize}

\textbf{Protocol.}
\begin{enumerate}
\item \textbf{Attractor topology:} Persistent homology $\to$ Betti curves $(\beta_0(r), \beta_1(r))$; topological analysis.
\item \textbf{Hop detector:} Define hop as $(|\Delta\beta_1| > \theta)$ within 1 s.
\item \textbf{Complexity scan:} White noise $\to$ music $\to$ speech $\to$ pure tone (5 levels).
\item \textbf{Permutation test:} For each configuration, shuffle timestamps $10^6$ times $\to$ p-value.
\item \textbf{Mismatch control:} Change geometry (10\%, 20\%, 50\%) $\to$ measure $(r_{sync})$ vs. dissimilarity; prediction: $(r_{sync} \propto e^{-d_\sigma/\lambda_\sigma})$.
\end{enumerate}

\textbf{Goals:}
\begin{itemize}
\item $(r_{sync})$ maximized at intermediate complexity (music/speech),
\item Overrepresentation at match: $(p < 10^{-6})$,
\item $(r_{sync})$ falls monotonically with increasing mismatch.
\end{itemize}

\textbf{Analysis:}
\begin{itemize}
\item Log-linear fit gives $(\lambda_\sigma)$,
\item \textbf{FDR correction} for 5 complexity levels and mismatch levels.
\end{itemize}

\textbf{Inference:} Null result $\to (\lambda_\sigma)$-bound and $(\varepsilon Q < 10^{-15})$.

