\section{Experimental Protocols (No-Loophole) and Parameter Inference}

All three proposed experiments are designed to be as close as practicable to ``no-loophole'' tests in the spirit of modern Bell experiments. This means:
\begin{itemize}
  \item full pre-registration of hypotheses, data-taking procedures, and analysis pipelines (e.g.\ on the Open Science Framework);
  \item cryptographic commit--reveal schemes for any codebooks or analysis choices that could otherwise be tuned post hoc;
  \item strict spacelike separation of laboratories where FTL influence is claimed;
  \item aggressive use of control conditions, sham runs, and blinding.
\end{itemize}
In all cases, raw data and analysis code should be published openly.
In all cases the observables we propose to measure (bit error rates,
energy balances, correlation functions) are classical summary
statistics of underlying quantum dynamics; any positive signal would
indicate the presence of the additional $S$-mediated channel rather
than a breakdown of standard quantum measurement theory.

\textbf{Statistical methodology.} For each experiment we pre-register:
\begin{itemize}
  \item the primary test statistic (e.g.\ BER difference, energy imbalance, or correlation coefficient);
  \item the null and alternative hypotheses in terms of that statistic;
  \item a target $p$-value or Bayes factor for claiming evidence;
  \item a multiple-comparisons correction (e.g.\ Benjamini--Hochberg FDR) across pre-specified families of tests.
\end{itemize}
A separate calibration study, using only local controls, is used to validate analysis code and to estimate systematic uncertainties; that calibration is assigned its own DOI.

\subsection{E1. Neuromorphic Ansible (Information)}

\textbf{Setup.}
\begin{itemize}
\item Two photonic or electronic reservoirs (e.g.\ 3D random resistor--capacitor or memristive networks with $N\sim 10^3$ nodes), labelled $A$ and $B$.
\item Both reservoirs are trained on an identical dataset (e.g.\ MNIST images, speech, or video snippets), using local learning rules only.
\item Each laboratory is enclosed in a Faraday cage and optically isolated (fiber-air gaps, optical isolators). Power is supplied by batteries during runs.
\item Independent atomic clocks (GPS-disciplined or ultra-stable crystal/OCXO) with jitter $\lesssim 1$\,ns provide local timing.
\item Laboratories are separated by $>1$\,km, so that light-travel time between them exceeds the timing resolution ($>3\,\mu$s).
\end{itemize}

From a technological perspective, all ingredients required for E1
(photonic or electronic reservoirs with $N\sim 10^3$ nodes, high-quality
pumps, and synchronized clocks over kilometre baselines) are available
with current or near-term hardware.  The main challenge lies in
accumulating sufficient statistics and in enforcing strict spacelike
separation and analysis pre-registration, rather than in developing
entirely new device physics.  We therefore view E1 as a near-term
(5--10 year) experimental target.

\textbf{Protocol.}
\begin{enumerate}
\item \textbf{Commit--reveal.} Before data taking, a sender codebook (mapping message bits to reservoir drive patterns) and a schedule of send/idle windows are generated using a pseudo-random generator seeded by a quantum random number generator (QRNG). The hash (e.g.\ SHA-256) of the codebook and schedule is published.
\item \textbf{Distance ladder calibration.} A separate, blinded calibration run is performed in which the pattern similarity between the two reservoirs is modified in pre-specified steps (Sec.~7.3). This is used to fit or bound $\lambda_\sigma$ and validate that any observed correlations track $d_\sigma$ as expected.
\item \textbf{Main run.} During each active send window, the sender reservoir is driven according to the codebook; during sham windows it is held in a neutral state. The receiver is driven by its local noise source or neutral state and records its high-dimensional output at the agreed sampling rate.
\item \textbf{Blinded analysis at receiver.} The receiver laboratory, without access to sender logs, runs correlation analyses against all possible codewords consistent with the published hash and schedule. Only after the analysis is fixed are the sender logs revealed.
\item \textbf{Environmental veto.} Muon and RF detectors are deployed at both sites; runs with unusually high cosmic-ray or RF activity are flagged or discarded.
\end{enumerate}

\textbf{Analysis.}
\begin{itemize}
\item Compute the BER for decoding the sender's message from the receiver's reservoir state using only locally available information.
\item Compare the BER between matched-structure runs and control runs in which the reservoirs are trained on independent datasets.
\item Examine how the BER improvement (or correlation coefficient) scales with the ladder level $\ell$ in the calibration, testing for the predicted $\exp[-d_\sigma(\ell)/\lambda_\sigma]$ dependence.
\end{itemize}

\textbf{Inference.} A null result at sensitivity $R_{\rm bit}^{({\rm null})}$ constrains the product $\varepsilon\,\lambda_\sigma\,\mathcal Q$ via the specialization of Eq.~(8.3) to the E1 geometry. A positive result consistent with the ladder scaling would give a direct estimate of $\lambda_\sigma$ and, combined with the pump power budget, a lower bound on $\varepsilon\mathcal Q$.

\subsubsection{Worked example: concrete neuromorphic-RC protocol}

As a concrete implementation, one can use:

\begin{itemize}
\item \textbf{Hardware:} 3D random resistor--capacitor networks with $N\sim 10^3$ nodes, trained on identical MNIST or audio datasets.
\item \textbf{Isolation:} Triple-nested Faraday cages, optical isolation (fiber-air-gap + optical isolators), battery-powered operation during runs. GPS-disciplined atomic clocks with jitter $<1$\,ns.
\item \textbf{Distance ladder:} Test 6 calibration levels (Sec.~7.3) in a blind pilot study with separate DOI. Freeze $\lambda_\sigma,\ell_0$ before the main run.
\item \textbf{Delayed choice:} QRNG selects codebook and time schedule after the hash is published but before transmission.
\item \textbf{Sham blocks:} Allocate 20\% of time blocks as ``sender-off'' controls to test for false positives (ghost-block detector).
\item \textbf{Cosmic veto:} Deploy muon detector; reject any run with $N_\mu>10$\,m$^{-2}$\,s$^{-1}$.
\item \textbf{Target:} $\Delta{\rm BER}\sim 10^{-3}$ (matched vs.\ mismatched) with $p<10^{-6}$ after Holm--Bonferroni correction for the 6 distance levels.
\end{itemize}

Statistical analysis includes sequential probability ratio test (SPRT) for early stopping, permutation tests ($\ge 10^6$ permutations), and Bayes factors comparing $H_1$ (resonance) vs.\ $H_0$ (chance + leakage).

\subsection{E2. Rotating energy tunnel and sidereal check}

\textbf{Setup.}
\begin{itemize}
\item Two macroscopic reservoirs $A$ and $B$ (e.g.\ cryogenic thermal masses or phase-change materials) mounted on a rigid, rotatable platform of size $L_{\rm exp}\sim 1$\,m.
\item Each reservoir is coupled to its own high-stability calorimeter with sensitivity $\delta E\lesssim 10^{-26}$\,J over $10^3$\,s.
\item The internal microstructure of $A$ and $B$ is engineered to be structurally matched in $S$ according to the chosen $O_S$ (similar defect patterns, topology, \emph{etc.}), but with different orientations in $M$ with respect to the platform axis.
\item A controlled pump (e.g.\ electrical, optical, or mechanical) drives the reservoirs in synchrony, providing the power $P_{\rm pump}$ entering Eq.~(6.1).
\end{itemize}

\textbf{Protocol.}
\begin{enumerate}
\item Alternate long integration windows with the platform fixed at different azimuthal orientations, sampling different angles between the apparatus and the substrate preferred frame.
\item Log $\Delta E_A$ and $\Delta E_B$ in each window and check Eq.~(12.2) for consistency, as well as any persistent bias in $\Delta E_A - \Delta E_B$.
\item Run for a total duration of $\sim 10^7$\,s spread over several weeks to resolve sidereal modulation at frequency $\Omega_\oplus$.
\item Perform control runs with pumps off, with deliberately de-structured media (destroying the pattern relevant for $O_S$), and with the reservoirs interchanged.
\end{enumerate}

\textbf{Analysis and inference.}
\begin{itemize}
\item Fit the time series of inferred $\Psig$ to the form of Eq.~(12.3), extracting limits or estimates of the mean $\bar\Psig$ and relative amplitude $\mathcal A$.
\item Translate $\bar\Psig$ into bounds on $\varepsilon\mathcal Q$ using Eq.~(6.1), given $P_{\rm pump}$ and an estimate of $\tilde{\Delta\Phi}$.
\item Use the sidereal amplitude $A_{\rm sid}^{(\mathrm{mat})}$ to constrain $\epsmat\Qmat(\lsig/L_{\rm exp})$ via Eq.~(12.4), taking into account photon-sector bounds on $\epsgam\Qgam$.
\end{itemize}

\subsubsection{Worked example: cryogenic micro-calorimetry}

A concrete realization for maximal sensitivity:

\begin{itemize}
\item \textbf{Hardware:} Two identical superconducting microwave cavities or metamaterial structures with cavity quality factor $Q_{\rm cav}\sim 10^6$.
\item \textbf{Calorimetry:} Operate at $T\sim 10$\,mK; transition-edge sensors or SQUID-based thermometry with $\delta T\sim 0.1\,\mu$K, yielding energy sensitivity $\delta E\sim 10^{-26}$\,J over integration time $\tau\sim 10^3$\,s.
\item \textbf{Pump modulation:} Pump reservoir $A$ with $P_{\rm pump}\approx 1\,\mu$W in 100\,s on/off cycles; keep $B$ below threshold.
\item \textbf{Matching test:} Vary internal geometry to achieve 0\%, 50\%, 100\% structural match. Expect correlation coefficient $r>0.8$ in fully matched condition.
\item \textbf{Latency scan:} Correlate $\Delta T_B(t)$ with $P_A(t-\delta)$, scanning $\delta\in[-10\,\mu{\rm s}, +10\,\mu{\rm s}]$. An FTL signal should peak at $\delta<-3\,\mu{\rm s}$ (spacelike), while thermal leakage would have $\delta>0$.
\item \textbf{Gravitational sidekick:} Deploy torsion pendulum or optical cavity in $B$ with acceleration sensitivity $\sim 10^{-15}$\,m\,s$^{-2}$ to test momentum-neutrality (Eq.~3.7). Report expected $\Delta\Phi_g$ even if below detection threshold. Note that we do \emph{not} expect any detectable gravitational effect at the target parameter values; this measurement simply provides an independent consistency check on momentum transfer.
\item \textbf{Rotation test:} Mount on rotating platform (0.1\,rpm) for full $2\pi$ azimuthal scan over several weeks ($\sim 10^7$\,s total), resolving sidereal modulation at $\Omega_\oplus\approx 7.3\times10^{-5}$\,rad\,s$^{-1}$.
\end{itemize}

\textbf{Ambitious target:} $|\Delta E_A+\Delta E_B|>10^{-25}$\,J with sidereal amplitude $A_{\rm sid}\ge 10^{-20}$.

\textbf{Null bound:} If $|\Delta E|<10^{-26}$\,J after $10^6$\,s, this constrains $\varepsilon\omega_0\mathcal Q<10^{-8}$\,Hz. A null sidereal signal gives $\varepsilon\mathcal Q<10^{-20}$.

\subsection{E3. Complexity scan and optimal patterning}

\textbf{Setup.}
\begin{itemize}
\item A platform whose configurational complexity can be tuned systematically, e.g.\ an electronic network with adjustable connectivity, a photonic lattice with controlled disorder, or a spin system near a phase transition.
\item Two copies of the platform constructed so that, for each complexity setting, they are structurally matched in $S$ according to $O_S$.
\end{itemize}

\textbf{Protocol.}
\begin{enumerate}
\item Define a scalar complexity measure $\Sigma$ (e.g.\ spectral entropy, participation ratio, or algorithmic compressibility of typical states).
\item Scan $\Sigma$ over a wide range by changing the control parameters of the platform, ensuring that $\Sigma$ can be estimated locally at each site.
\item For each complexity setting, run an E1- or E2-type protocol to infer an effective hop rate $r_{\rm hop}(\Sigma)$ or an associated measure of FTL strength.
\item Include sham configurations in which the two copies are deliberately de-correlated in $S$, as controls.
\end{enumerate}

\textbf{Analysis and inference.}
\begin{itemize}
\item Test whether $r_{\rm hop}(\Sigma)$ is consistent with zero across all $\Sigma$ (null hypothesis) or exhibits a statistically significant peak near some $\Sigma_{\rm opt}$ (alternative).
\item If a peak is found, compare its width and location with theoretical expectations for the complexity dependence of the pattern quality factor $\mathcal Q(\Sigma)$.
\item A null result constrains $\lambda_\sigma$ and $\mathcal Q$ for the explored class of patterns, complementing the point constraints from E1 and E2.
\end{itemize}

In practice, realizing a clean scan over configurational entropy $\Sigma$
with precise control over disorder and pump parameters is challenging,
especially in platforms with many coupled degrees of freedom.  We view
E3-type tests as longer-term goals that may require one or two orders of
magnitude further progress in controllable complex systems beyond what
is currently available.

\subsubsection{Worked example: chaos-to-chaos synchronization test}

A concrete realization using turbulent flows:

\begin{itemize}
\item \textbf{Hardware:} Two Rayleigh--B\'enard convection cells or reaction--diffusion systems ($L\sim 10$\,cm) with identical geometry. Monitor via laser Doppler velocimetry or imaging at 1\,kHz sampling rate.
\item \textbf{Drive:} Modulate heat flux or chemical concentration with variable complexity: white noise $\to$ music $\to$ speech $\to$ pure tone (5 levels).
\item \textbf{Attractor topology:} Use persistent homology to compute Betti curves $(\beta_0(r),\beta_1(r))$ from the time-delay embedding of velocity/concentration fields.
\item \textbf{Hop detector:} Define a ``hop'' event as $|\Delta\beta_1|>\theta$ within 1\,s; look for temporal correlations in hop events between the two systems.
\item \textbf{Mismatch control:} Systematically vary geometry (10\%, 20\%, 50\% dissimilarity) and measure synchronization rate $r_{\rm sync}$. Prediction: $r_{\rm sync}\propto\exp[-d_\sigma/\lambda_\sigma]$.
\item \textbf{Permutation test:} For each complexity level and mismatch setting, shuffle timestamps $10^6$ times to compute $p$-value for observed correlations.
\item \textbf{Statistical correction:} Apply Benjamini--Hochberg FDR correction for 5 complexity levels and multiple mismatch settings.
\end{itemize}

\textbf{Target:} Synchronization rate $r_{\rm sync}$ maximized at intermediate complexity (music/speech), with overrepresentation at match giving $p<10^{-6}$ (corrected). The rate should fall monotonically with increasing structural mismatch.

\textbf{Inference:} A log-linear fit of $r_{\rm sync}$ vs.\ mismatch yields an estimate of $\lambda_\sigma$. A null result constrains $\varepsilon\mathcal Q<10^{-15}$.
