\section{Predictions and Numerical Targets}

\textbf{Negative predictions (should not be seen):}

\begin{itemize}
\item No deviations in gravitational laws, vacuum dispersion, or torsion wave experiments.
\item No robust effects in homogeneous crystals $(N \gg 1 \to P_\sigma \to 0)$.
\item No signals in accelerator experiments $(O_S$ irrelevant, $\Delta > 4$).
\item No everyday signaling without shared structure and pump.
\end{itemize}

\textbf{Positive predictions with numerical targets:}

\subsection{Pred. 1: Twin-Reservoir Correlations (E1)}

\textbf{Goal:} Bit error rate (BER) scales as:
\begin{equation}
BER = \frac{1}{2} \left( 1 - \betaBER \, e^{-d_\sigma/\lambda_\sigma} \right).
\end{equation}

For $(d_\sigma = 0)$ (identical dataset): $(BER_{min} = (1-\betaBER)/2 \approx 0.25)$ (for $\betaBER \sim 0.5$).

For $(d_\sigma \gg \lambda_\sigma)$ (mismatch): $(BER \to 0.5)$ (chance).

\textbf{Numerical target (adjusted):}
\begin{itemize}
\item Report \textbf{$\Delta$BER} at \textbf{$\sim10^{-3}$}-level (match vs. mismatch) with $10^9$ bits,
\item primary test statistic: \textbf{cross-correlation/coherence} (not hard BER thresholds),
\item analysis: \textbf{SPRT}, permutation tests, Holm-Bonferroni.
\item \textbf{Delayed choice} (QRNG) and spacelike separation satisfied.
\end{itemize}

\textbf{Null bound:} If $(BER \geq 0.49)$ for all configurations $\to (\varepsilon \lambda_\sigma Q < 10^{-12}~\text{m})$.

\subsection{Pred. 2: Energy Tunnel (E2)}

\textbf{Goal:} Differential energy balance:
\begin{equation}
\Delta E_A + \Delta E_B = \Psig \cdot \Delta t,
\tag{12.2}
\end{equation}

where $(\Psig \sim \varepsilon \hbar\omega_0 \Ksig \mathcal Q \tilde{\mathcal K} \tilde{\Delta\Phi})$.

\textbf{Three scenarios for E2 detectability:}

\begin{table}[h]
\centering
\begin{tabular}{llll}
\toprule
Scenario & Q & $\Delta E$ (10$^3$ s) & Detectability \\
\midrule
\textbf{Baseline} & $10^{-5}$ & $\sim10^{-40}$ J & Not detectable ($\delta E \sim 10^{-26}$ J) \\
\textbf{Target} & $10^{-3}$ & $\sim10^{-27}$ J & Below limit but approaching \\
\textbf{Ambitious} & $10^{-2}$ & $\sim10^{-25}$ J & Marginally detectable at limit \\
\bottomrule
\end{tabular}
\end{table}

With power form (6.1), $\Psig=\varepsilon \Ksig \mathcal Q \tilde{\Delta\Phi} P_{\rm pump}$.

\textbf{Parameter example (``target'' near sensitivity limit):}
\begin{itemize}
\item $P_{\rm pump}=1\,\mu\mathrm{W}$, $\Ksig\simeq0.5$, $\tilde{\Delta\Phi}\simeq1$,
\item $\varepsilon=10^{-15}$, $\mathcal Q=10^{-8}$ (compatible with (11.4)),
\item[] $\to \Psig\approx 5\times10^{-30}\,\mathrm{W}$ $\Rightarrow \Delta E(10^3\mathrm{~s})\approx 5\times10^{-27}\,\mathrm{J}$.
\end{itemize}

This lies just below $\delta E\sim10^{-26}\,\mathrm{J}$ (mK calorimetry) and indicates realistic target level.

\begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,title=\textbf{Worked Numerical Examples}]
\textbf{E2 Energy Tunnel – Target Parameter Set:}

Given: $P_{\rm pump} = 1\,\mu{\rm W}$, $\Ksig = 0.5$, $\tilde{\Delta\Phi} \simeq 1$, $\varepsilon = 10^{-15}$, $\mathcal{Q} = 10^{-8}$.

Using power formula (6.1): $\Psig = \varepsilon \Ksig \mathcal{Q} \tilde{\Delta\Phi} \, P_{\rm pump}$:
\begin{align*}
\Psig &= 10^{-15} \times 0.5 \times 10^{-8} \times 1 \times 10^{-6}\,{\rm W} \\
      &= 5 \times 10^{-30}\,{\rm W}.
\end{align*}

Over measurement window $\Delta t = 10^3$ s:
\[
\Delta E = \Psig \cdot \Delta t = 5\times10^{-30}\,{\rm W} \times 10^{3}\,{\rm s} = 5\times10^{-27}\,{\rm J}.
\]

\textbf{Detectability:} mK calorimetry sensitivity $\delta E \sim k_B T \sim 10^{-26}$ J at $T = 1$ mK. Therefore $\Delta E/\delta E \sim 0.5$ (edge of sensitivity; requires long integration or lower $T$).

\medskip

\textbf{E2 Sidereal Modulation – Cross-Sector Check:}

For rotation test, sidereal amplitude (matter sector, Eq. 12.4):
\[
A_{\rm sid}^{(\rm mat)} \simeq \epsmat \left(\frac{\lsig}{L_{\rm exp}}\right) \Qmat \, \Xi.
\]

Target sensitivity: $A_{\rm sid} \gtrsim 10^{-20}$ over $10^7$ s.

With $L_{\rm exp} \sim 1$ m, $\lsig \sim 10^{-6}$ m, $\Xi \sim 1$:
\[
A_{\rm sid} \sim \epsmat \times 10^{-6} \times \Qmat.
\]

Requiring $A_{\rm sid} \gtrsim 10^{-20}$ gives:
\[
\epsmat \, \Qmat \gtrsim 10^{-14}.
\]

\textbf{Cross-check:} Photon-sector bound (11.4) constrains $\epsgam \Qgam \lesssim 10^{-24}$. If matter and photon sectors have similar couplings ($\epsmat \sim \epsgam \sim 10^{-15}$), then target $\Qmat \sim 10^{1}$ (unitless) is \emph{nine orders of magnitude} larger than allowed photon $\Qgam \sim 10^{-9}$—consistent with sector separation hypothesis (see Appendix F).
\end{tcolorbox}

\textbf{Detection limit:} Micro-calorimetry at mK level: $(\delta E \sim k_B T \sim 10^{-26}~\text{J})$ (for $T \sim 1$ mK) with cryo technology.

\textbf{Microscopic anti-correlation:} energy balance between A/B.

\textbf{Null bound:} If no signal after $10^6$ s with $(\delta E = 10^{-26}~\text{J})$ $\to (\varepsilon \omega_0 Q < 10^{-8}~\text{Hz})$.

\subsection{Pred. 3: Anisotropic Daily Modulation (E2 – Rotation Test; \textbf{Matter Sector})}

\textbf{Goal:} Power modulation with period 23 h 56 min (sidereal):
\begin{equation}
  \Psig(t) \;=\; \bar \Psig\left[1+ \mathcal A\cos\!\big(\Omega_\oplus t + \phi\big)\right],
  \tag{12.3}
\end{equation}
with amplitude \textbf{in the matter sector}
\begin{equation}
  A_{\rm sid}^{(\mathrm{mat})} \;\simeq\;
  \frac{\epsmat}{\epsgam} \cdot \frac{\lsig}{L_{\rm exp}} \cdot \frac{\Qmat}{\Qgam} \cdot \Xi \,,
  \tag{12.4}
\end{equation}
\textbf{Important:} The prediction here concerns the \textbf{matter sector} $(\epsmat, \Qmat)$. Bounds in §11.4 constrain $\epsgam\Qgam$ (optics) and are therefore not directly binding on matter-sector measurements; see Appendix F (sector separation).

\textbf{Cross-reference:} Report $\hat{A}_{\rm sid}$ and compare against (11.4) and (12.4); specify which measurement track (optical vs. material) is used in the rig.

\textbf{Numerical target (3$\sigma$, $10^7$ s):}
\begin{equation}
A_{\rm sid}\;\gtrsim\;10^{-20}.
\end{equation}
\textbf{Stretch:} $A_{\rm sid}\gtrsim5\times10^{-21}$.

\textbf{Null bound:} If $\hat{A}_{\rm sid}<10^{-20}$ obtain $\epsmat\Qmat<10^{-20}$ for given $L_{\rm exp}$ and $\lsig$ (report both).

\subsection{Pred. 4: Complexity Optimum (E3)}

\textbf{Goal:} Sync-hop rate $(r_{sync})$ vs. drive complexity $(\Sigma_{drive})$:

\begin{equation}
r_{sync} = r_0 \, \Sigma_{drive} \, e^{-\Sigma_{drive} / \Sigma_{opt}},
\end{equation}

maximum at $(\Sigma_{opt})$ (rich but compressible).

\textbf{Numerical target:}
\begin{itemize}
\item White noise $(\Sigma \sim \infty)$: $(r_{sync} \sim 0)$,
\item Pure tone $(\Sigma \sim 0)$: $(r_{sync} \sim 0)$,
\item Music/speech $(\Sigma \sim \Sigma_{opt} \sim 5$ bits/sample$)$: $(r_{sync} \sim \max)$.
\end{itemize}

\textbf{Test:} Vary from pure sine wave $\to$ complex sound $\to$ noise; measure $(r_{sync})$; expect unimodal curve.

\textbf{Excess:} simultaneous attractor hops when drive rhymes (small $(d_\sigma)$); falls away at mismatch.

\subsection{Summary Predictions $\to$ Bounds}

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
Experiment & Positive signal & Null bound \\
\midrule
E1 (ansible) & $\Delta$BER $\sim 10^{-3}$ (match vs. mismatch), cross-corr. & $\varepsilon \lambda_\sigma Q < 10^{-12}$ m \\
E2 (energy) & $\Delta E > 10^{-25}$ J $(Q \sim 10^{-2})$ & $\varepsilon \omega_0 Q < 10^{-8}$ Hz \\
E2 (rotation, \textbf{matter}) & $A_{sid} \geq 10^{-20}$ (3$\sigma$, $10^7$ s) & $\varepsilon_{mat} Q_{mat} < 10^{-20}$ \\
E3 (chaos) & $r_{sync}$ peak at $\Sigma_{opt}$ & $\varepsilon Q < 10^{-15}$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{End-to-end parameter map}

\textbf{Which observables constrain which parameter products (at leading order):}

\begin{table}[h]
\centering
\small
\begin{tabular}{lll}
\toprule
Observable & Primary constraint & Secondary/notes \\
\midrule
E1 ($\Delta$BER, coherence) & $\varepsilon\,\lambda_\sigma\,\mathcal Q$ (via (8.3–8.4)) & Distance ladder (§7.3) \\
E2 ($\Delta E$ over time) & $\varepsilon\,\omega_0\,\mathcal Q$ & Power form (6.1) \\
E2 (sidereal amplitude) & $\varepsilon\,\mathcal Q\cdot(\lambda_\sigma/L_{\rm exp})$ & Geometry factor $\Xi$ (11.2) \\
E3 (hop rate vs. complexity) & $\lambda_\sigma$, $\mathcal Q$ & Peak at $\Sigma_{opt}$ \\
\bottomrule
\end{tabular}
\end{table}

