\section{Worked Experimental Protocols (Illustrative)}

\subsection{Neuromorphic ansible (E1) – detailed protocol}

As a concrete implementation, one can use:

\begin{itemize}
\item \textbf{Hardware:} 3D random resistor--capacitor networks with $N\sim 10^3$ nodes, trained on identical MNIST or audio datasets.
\item \textbf{Isolation:} Triple-nested Faraday cages, optical isolation (fiber-air-gap + optical isolators), battery-powered operation during runs. GPS-disciplined atomic clocks with jitter $<1$\,ns.
\item \textbf{Distance ladder:} Test 6 calibration levels (Sec.~7.3) in a blind pilot study with separate DOI. Freeze $\lambda_\sigma,\ell_0$ before the main run.
\item \textbf{Delayed choice:} QRNG selects codebook and time schedule after the hash is published but before transmission.
\item \textbf{Sham blocks:} Allocate 20\% of time blocks as ``sender-off'' controls to test for false positives (ghost-block detector).
\item \textbf{Cosmic veto:} Deploy muon detector; reject any run with $N_\mu>10$\,m$^{-2}$\,s$^{-1}$.
\item \textbf{Target:} $\Delta{\rm BER}\sim 10^{-3}$ (matched vs.\ mismatched) with $p<10^{-6}$ after Holm--Bonferroni correction for the 6 distance levels.
\end{itemize}

Statistical analysis includes sequential probability ratio test (SPRT) for early stopping, permutation tests ($\ge 10^6$ permutations), and Bayes factors comparing $H_1$ (resonance) vs.\ $H_0$ (chance + leakage).

\subsection{Cryogenic micro-calorimetry (E2) – detailed protocol}

A concrete realization for maximal sensitivity:

\begin{itemize}
\item \textbf{Hardware:} Two identical superconducting microwave cavities or metamaterial structures with cavity quality factor $Q_{\rm cav}\sim 10^6$.
\item \textbf{Calorimetry:} Operate at $T\sim 10$\,mK; transition-edge sensors or SQUID-based thermometry with $\delta T\sim 0.1\,\mu$K, yielding energy sensitivity $\delta E\sim 10^{-26}$\,J over integration time $\tau\sim 10^3$\,s.
\item \textbf{Pump modulation:} Pump reservoir $A$ with $P_{\rm pump}\approx 1\,\mu$W in 100\,s on/off cycles; keep $B$ below threshold.
\item \textbf{Matching test:} Vary internal geometry to achieve 0\%, 50\%, 100\% structural match. Expect correlation coefficient $r>0.8$ in fully matched condition.
\item \textbf{Latency scan:} Correlate $\Delta T_B(t)$ with $P_A(t-\delta)$, scanning $\delta\in[-10\,\mu{\rm s}, +10\,\mu{\rm s}]$. An FTL signal should peak at $\delta<-3\,\mu{\rm s}$ (spacelike), while thermal leakage would have $\delta>0$.
\item \textbf{Gravitational sidekick:} Deploy torsion pendulum or optical cavity in $B$ with acceleration sensitivity $\sim 10^{-15}$\,m\,s$^{-2}$ to test momentum-neutrality (Eq.~3.7). Report expected $\Delta\Phi_g$ even if below detection threshold. Note that we do \emph{not} expect any detectable gravitational effect at the target parameter values; this measurement simply provides an independent consistency check on momentum transfer.
\item \textbf{Rotation test:} Mount on rotating platform (0.1\,rpm) for full $2\pi$ azimuthal scan over several weeks ($\sim 10^7$\,s total), resolving sidereal modulation at $\Omega_\oplus\approx 7.3\times10^{-5}$\,rad\,s$^{-1}$.
\end{itemize}

\textbf{Ambitious target:} $|\Delta E_A+\Delta E_B|>10^{-25}$\,J with sidereal amplitude $A_{\rm sid}\ge 10^{-20}$.

\textbf{Null bound:} If $|\Delta E|<10^{-26}$\,J after $10^6$\,s, this constrains $\varepsilon\omega_0\mathcal Q<10^{-8}$\,Hz. A null sidereal signal gives $\varepsilon\mathcal Q<10^{-20}$.

\textbf{Preliminary Noise Budget (Target $\delta E \sim 10^{-26}$ J):}
To achieve the target sensitivity, the following noise sources must be mitigated:
\begin{itemize}
\item \textbf{Thermal Fluctuation Noise (Phonon Noise):} $\Delta E_{th} = \sqrt{k_B T^2 C}$. For a $1$g silicon absorber at $10$ mK, $\Delta E_{th} \approx 10^{-19}$ J. \textbf{Mitigation:} Requires mass reduction to $\mu$g scale (MEMS calorimeters) or non-equilibrium inductive sensing (KES).
\item \textbf{Vibrational Heating:} Micro-fracture events in suspension. \textbf{Mitigation:} Underground facility with active vibration isolation stages ($>60$ dB suppression).
\item \textbf{Cosmic Ray Muons:} A single muon deposits $\sim$MeV ($10^{-13}$ J). \textbf{Mitigation:} Deep underground location ($>2$ km.w.e.) and active muon veto (scintillator panels) to reject coincidence events.
\item \textbf{Two-Level System (TLS) Noise:} Dielectric loss in oxide layers. \textbf{Mitigation:} Crystalline sapphire substrates and trenching to remove surface oxides.
\end{itemize}

\emph{Assessment:} The "Ambitious" target requires integrating state-of-the-art dark matter detector technology (e.g., from SuperCDMS/CRESST) into a rotating setup, confirming this is a Phase 3 decadal challenge.

\subsection{Chaos-to-chaos synchronization test (E3) – detailed protocol}

A concrete realization using turbulent flows:

\begin{itemize}
\item \textbf{Hardware:} Two Rayleigh--B\'enard convection cells or reaction--diffusion systems ($L\sim 10$\,cm) with identical geometry. Monitor via laser Doppler velocimetry or imaging at 1\,kHz sampling rate.
\item \textbf{Drive:} Modulate heat flux or chemical concentration with variable complexity: white noise $\to$ music $\to$ speech $\to$ pure tone (5 levels).
\item \textbf{Attractor topology:} Use persistent homology to compute Betti curves $(\beta_0(r),\beta_1(r))$ from the time-delay embedding of velocity/concentration fields.
\item \textbf{Hop detector:} Define a ``hop'' event as $|\Delta\beta_1|>\theta$ within 1\,s; look for temporal correlations in hop events between the two systems.
\item \textbf{Mismatch control:} Systematically vary geometry (10\%, 20\%, 50\% dissimilarity) and measure synchronization rate $r_{\rm sync}$. Prediction: $r_{\rm sync}\propto\exp[-d_\sigma/\lambda_\sigma]$.
\item \textbf{Permutation test:} For each complexity level and mismatch setting, shuffle timestamps $10^6$ times to compute $p$-value for observed correlations.
\item \textbf{Statistical correction:} Apply Benjamini--Hochberg FDR correction for 5 complexity levels and multiple mismatch settings.
\end{itemize}

\textbf{Target:} Synchronization rate $r_{\rm sync}$ maximized at intermediate complexity (music/speech), with overrepresentation at match giving $p<10^{-6}$ (corrected). The rate should fall monotonically with increasing structural mismatch.

\textbf{Inference:} A log-linear fit of $r_{\rm sync}$ vs.\ mismatch yields an estimate of $\lambda_\sigma$. A null result constrains $\varepsilon\mathcal Q<10^{-15}$.
