\section{Causality – Formal Proof}

\textbf{Monotonic-$(T)$ Axiom:} All substrate-mediated transfers obey:
\begin{equation}
  \Delta\tau>0\ \ \text{for all substrate transfers},\qquad
  \text{all kernels are retarded in }T.
  \tag{10.1}
\end{equation}

\textbf{No causal loops:} All substrate-mediated influences are retarded in $(T)$ and strictly monotone with respect to the khronon $(\tau)$ (Eq. (10.1)). Any observable influence in $(M)$ is a concatenation of (i) $(M)$-causal propagation with respect to $(g_{\mu\nu})$ and (ii) $(S)$-local transfers retarded in $(T)$. Since $(T)$ induces a partial order on the combined interaction graph $(\mathcal G_M \cup \mathcal G_S)$, and edges of $(\mathcal G_S)$ are directed by retardedness, closed $(T)$-nonincreasing loops are forbidden. Therefore anti-telephone constructions fail even for relatively moving observers.

\begin{theorem}[Causal monotonicity]
Under conditions (i) all resonance dynamics is retarded in substrate ordering $(T)$, (ii) each resonance step requires $\tilde{\mathcal K}(e) \ge 0$, and (iii) $\varepsilon$ is finite, there exist no closed causal loops in $(M\times S)$.
\end{theorem}

\begin{tcolorbox}[colback=green!5!white,colframe=green!75!black,title=\textbf{Formal Categorical Causality Proof}]
\textbf{Construction:} Define category $\mathcal{C}$ of substrate events:
\begin{itemize}[leftmargin=*,noitemsep,topsep=3pt]
\item \textbf{Objects:} States $s \in M \times S$ with spacetime location and substrate configuration.
\item \textbf{Morphisms:} Allowed transitions $\phi: s \to s'$ (local M-updates or S-resonance).
\end{itemize}

\textbf{Partially ordered set:} Equip $(\mathcal{C}, \prec)$ with ordering $s \prec s'$ iff there exists causal path $s \to s'$.

\textbf{Time functor:} Define $T: \mathcal{C} \to (\mathbb{N}, \leq)$ mapping each state to its substrate tick count $T(s) \in \mathbb{N}$. For any morphism $\phi: s \to s'$:
\[
T(s') > T(s) \quad \text{(strict monotonicity)}.
\]

\textbf{Functor properties:}
\begin{enumerate}[leftmargin=*,noitemsep,topsep=3pt]
\item $T$ preserves composition: $T(\phi_2 \circ \phi_1) = T(\phi_2)$ with intermediate $T$ values strictly increasing.
\item Each morphism has non-negative cost: $\tilde{\mathcal{K}}(\phi) \ge 0$.
\item Composite cost is additive: $\tilde{\mathcal{K}}(\phi_2 \circ \phi_1) = \tilde{\mathcal{K}}(\phi_1) + \tilde{\mathcal{K}}(\phi_2)$.
\end{enumerate}

\textbf{Loop impossibility:} Suppose closed loop $L: s_0 \to s_1 \to \cdots \to s_n = s_0$. Then:
\[
T(s_n) = T(s_0) \quad \text{but} \quad T(s_n) > T(s_{n-1}) > \cdots > T(s_0) \quad \Rightarrow\Leftarrow
\]
Contradiction. Therefore no closed loops exist in category with strictly monotone time functor.

\textbf{Conclusion:} The substrate's discrete update ordering $T$ provides an \emph{absolute arrow of time} preventing causal loops even when spacetime transfers appear FTL. \qed
\end{tcolorbox}

\textbf{Proof (sketch).}

\begin{enumerate}
\item \textbf{Category construction.} Let $(\mathcal{C})$ be the category of substrate events, where objects are states at different $(T)$ and morphisms are update steps (either local in M or resonance via S).

\item \textbf{Time functor.} Define $(T: \mathcal{C} \to (\mathbb{N}, \leq))$, $(T(s_i) = T_i)$. Require each morphism $(f: s \to s')$ has $(T(s') > T(s))$ (strict monotonicity).

\item \textbf{Cost monotonicity.} Each resonance morphism $(\phi: s \to s')$ requires $\tilde{\mathcal K}(\phi)\ge0$. Composed morphism $(\phi_2 \circ \phi_1)$ has $\tilde{\mathcal K}(\phi_2\circ\phi_1)=\tilde{\mathcal K}(\phi_1)+\tilde{\mathcal K}(\phi_2)$.

\item \textbf{Loop analysis.} Assume closed loop in M: events $(A \to B \to C \to A)$ with spacetime coordinates such that final step is in A's past light cone. This requires a morphism chain with $(\Delta T \leq 0)$ or $\tilde{\mathcal K}_{\rm total}<0$. Both are forbidden by construction.

\item \textbf{Multi-frame test.} In arbitrary Lorentz frame, M-coordinates transform, but $(T)$ is frame-invariant (substrate-absolute). Loop in one frame = loop in all frames $\to$ forbidden.
\end{enumerate}

\textbf{Conclusion:} No antitelephones. FTL in M is compensated by T-monotonicity in substrate. \qed

\textbf{Rule (anti-telephone):} Resonance transfers are \emph{only} permitted if
\begin{equation}
\Delta\tau>0
\end{equation}
in substrate time along each substep in the chain. The rule is locally testable (apparatus phase-locking to $u^\mu$) and forbids causal loops even for parties in relative motion.

Complete category-theoretic proof is in Appendix D.

\begin{corollary}[Two-lab anti-telephone]
For any two laboratories with four-velocities $U_A^\mu$ and $U_B^\mu$, any chain of resonance steps that satisfies $\Delta\tau>0$ on \textbf{each} substep in the local control logic of each apparatus cannot form a causal loop. The condition is frame-independent because $\tau$ is a scalar and increases monotonically along allowed morphisms.
\end{corollary}

