\section{Nomenclature and Notation}
\label{app:nomenclature}

\begin{itemize}
  \item \textbf{Aether resonance:} substrate-local coupling in pattern space $S$ that, when projected to spacetime $M$, appears as superluminal transfer.

  \item \textbf{Pattern space $S$:} space of local data structures; the distance $d_\sigma$ measures (approximate) isomorphic similarity of patterns and carries dimensions of length (meters) via the embedding scale $\ell_0$.

  \item $\varepsilon$: dimensionless microscopic coupling strength between visible sector and substrate (benchmark size $\sim 10^{-15}$ unless otherwise stated).

  \item $\alpha$: gravitational coupling factor multiplying $T^S_{\mu\nu}$ in the Einstein equations. Consistency with the Bianchi identity in the presence of the exchange current $J^\nu_\sigma$ forces a \emph{universal} coupling
  \[
    \alpha \equiv 1
  \]
  as shown in Sec.~\ref{subsec:alpha-constraint}; there is no tunable ``reduced'' gravitational coupling for the $S$-sector.

  \item $\hbar\omega_0$: characteristic energy scale of the driven transition in a given experiment (e.g.\ $\hbar\omega_0 \sim 10^{-23}\,\mathrm{J}$ at 10\,GHz).

  \item $\lambda_\sigma$: structural coherence/range in $S$; enters the similarity kernel $K_\sigma = \exp[-d_\sigma/\lambda_\sigma]$ and has units of length (meters).

  \item $\ell_0$: embedding scale used in the definition of $d_\sigma$, setting the typical size of structural changes that correspond to $d_\sigma\sim 1$ (benchmark $\ell_0 \sim 1\,\mu\mathrm{m}$).

  \item $Q$ (or $\mathcal{Q}$): pattern coherence/quality factor. Dimensionless, with
  \[
    0 \le Q \le 1,
  \]
  where $Q\sim 10^{-10}$ corresponds to typical disordered or thermal configurations and $Q\sim 10^{-2}$ is an ambitious target for carefully engineered, near-critical structures.

  \item $Q_{\mathrm{cav}}$: electromagnetic cavity quality factor in E2-type setups (Sec.~13.2.1). Dimensionless, characterizing the photon storage time and energy decay rate in the optical or microwave resonator.

  \item $K_\sigma$: similarity kernel on $S$,
  \[
    K_\sigma(\sigma,\sigma') = \exp[-d_\sigma(\sigma,\sigma')/\lambda_\sigma],
  \]
  dimensionless and positive semidefinite.

  \item $\Gamma_{\mathrm{pump}}$: pump-rate (or cost) factor, defined by
  \[
    \Gamma_{\mathrm{pump}} = \frac{P_{\mathrm{pump}}}{\hbar\omega_0} \quad [\mathrm{s}^{-1}],
  \]
  where $P_{\mathrm{pump}}$ is the physical pump power.

  \item $O_S$: selection operator constructed from visible fields and curvature, with mass dimension $\Delta>4$ in its seed form and normalized to $[O_S]=4$ after division by a UV scale $\Lambda^{\Delta-4}$ (Sec.~5).

  \item $J^\nu_\sigma$: exchange four-current between visible and $S$-sector, with dimensions of energy density per unit time,
  \[
    [J^\nu_\sigma] = \mathrm{energy}\cdot\mathrm{length}^{-3}\cdot\mathrm{time}^{-1}.
  \]

  \item $T^S_{\mu\nu}$: energy-momentum tensor for the $S$-sector, including the aether/khronon contribution and any effective stress--energy induced by the mediator.

  \item $\chi(\sigma,T)$: $S$-mediator field propagating locally in pattern space $S$ at finite substrate speed $c_S$ and mediating resonance, with range $\lambda_\sigma = c_S/m_\chi$.

  \item \textbf{Degeneracy dilution:} suppression $\sim 1/N$ of the effective pattern signal when many macroscopically equivalent matches compete; in homogeneous/thermal states with $N\to\infty$ this drives $\langle O_S\rangle\to 0$.

  \item $\tau$: foliation scalar (``khronon'') defining the substrate’s preferred time function and providing the global ordering used in the causality proof.

  \item $u^\mu$: unit timelike vector field aligned with the preferred frame, $u^\mu u_\mu=-1$; appears in the aether/khronon Lagrangian and defines the foliation underlying $\tau$.
\end{itemize}
