\section{Discussion: What This Means}

\subsection{Scientific Implications}

\textbf{If the theory is right:}

\textbf{1. Spacetime is emergent, not fundamental:}\\
We've always suspected (quantum gravity, AdS/CFT) but never had evidence. Aether resonance would be direct experimental evidence of the substrate layer.

\textbf{2. New communication technology:}\\
FTL communication under controlled conditions. Not sci-fi ``ansible'' (unlimited bandwidth, instant messaging) but:
\begin{itemize}
\item Low bandwidth ($\sim$bits per minute in best case)
\item Requires matched structures
\item Thermodynamically costly
\item But: Truly non-local, no signal propagation
\end{itemize}

\textbf{Applications:}
\begin{itemize}
\item Secure communication (no eavesdropping - the signal doesn't travel through space!)
\item Deep space probes (no light-speed delay)
\item Fundamental physics experiments (probe substrate structure)
\end{itemize}

\textbf{3. Unified framework:}\\
Connects:
\begin{itemize}
\item Quantum mechanics (substrate as quantum cellular automaton?)
\item Relativity (emergent from coarse-graining)
\item Thermodynamics (substrate entropy and free energy)
\item Complexity theory (algorithmic similarity as physical resource)
\end{itemize}

\textbf{If the theory is wrong:}

\textbf{1. Stringent bounds on parameter space:}\\
Even null results are valuable:
\begin{itemize}
\item $\varepsilon \lambda_\sigma \mathcal{Q} < 10^{-12}$ m (from E1)
\item $\varepsilon \omega_0 \mathcal{Q} < 10^{-8}$ Hz (from E2)
\item $\varepsilon \mathcal{Q} < 10^{-20}$ (from anisotropy)
\end{itemize}

These constrain \emph{any} substrate-coupling theory, not just ours.

\textbf{2. Rules out a class of models:}\\
Any discrete substrate model with:
\begin{itemize}
\item Pattern-space coupling
\item Emergent relativity
\end{itemize}
Must satisfy our bounds or be excluded.

\textbf{3. Sharpens quantum foundations:}\\
Lieb-Robinson violations are tightly constrained. This informs quantum information theory, quantum computing (limits on non-local gates), and quantum gravity (holography bounds).

\subsection{Philosophical Implications}

\textbf{On the nature of reality:}

If spacetime is emergent, then ``what exists fundamentally?'' is deeper than we thought.

\emph{Analogy:} Asking ``what exists in Minecraft?'' has two answers:
\begin{itemize}
\item \textbf{Player level:} Blocks, mobs, redstone circuits (emergent objects)
\item \textbf{Code level:} Data structures, update loops, voxel arrays (substrate)
\end{itemize}

Both are real, but one is more fundamental.

\textbf{On causality:}

We tend to think causality = light-cone structure. But perhaps:
\begin{itemize}
\item \textbf{Emergent causality} (light cones) is what we see
\item \textbf{Substrate causality} ($T$-ordering) is what's fundamental
\end{itemize}

They usually agree, but aether resonance would show where they diverge.

\textbf{On information:}

If information can ``travel'' through substrate ($d_\sigma$) rather than spacetime ($|x-y|$), then:
\begin{itemize}
\item Information is more fundamental than energy-momentum
\item Structural similarity is a physical resource (like energy or entropy)
\item The universe ``computes'' at a deeper level than we observe
\end{itemize}

This connects to:
\begin{itemize}
\item ``It from bit'' (Wheeler)
\item Digital physics (Zuse, Wolfram, 't Hooft)
\item Computational universe hypothesis
\end{itemize}

\subsection{Why This Is Worth Testing}

Even with tiny signals and huge challenges, aether resonance deserves experimental attention because:

\textbf{1. Falsifiability:}\\
Clear predictions, null results $\to$ bounds. This is how science works.

\textbf{2. Asymmetric payoff:}
\begin{itemize}
\item \textbf{Null result:} Useful bounds, still publishable
\item \textbf{Positive result:} Revolutionary, Nobel-worthy
\end{itemize}

Risk/reward ratio favors trying.

\textbf{3. Technological spinoffs:}\\
Even if aether resonance doesn't exist:
\begin{itemize}
\item Developing cryo-calorimetry at $10^{-26}$ J sensitivity
\item Precision control of quantum states for high $\mathcal{Q}$
\item Topological data analysis for complex systems
\item Blind analysis protocols
\end{itemize}

All have independent value.

\textbf{4. Fundamental question:}\\
``Is spacetime fundamental or emergent?'' is one of the deepest open questions in physics. Any experimental handle, however indirect, is precious.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

