\section{Operationalized $d_\sigma$-Metric}

Algorithmic similarity is non-computable. We specify a \textbf{practical proxy}:

\textbf{Operational surrogate for algorithmic similarity:} We approximate "algorithmic isomorphism" by a finite feature map $\Phi(\sigma)=(\mathcal I_1,\ldots,\mathcal I_{N_{\rm feat}})$ of local gauge/diffeo invariants. We choose weights $(w_I>0)$, fix an embedding scale $(\ell_0)$, and define:

\begin{equation}
\begin{split}
  d_\sigma(\sigma,\sigma') &:= \ell_0\,
  \left(\sum_{I=1}^{N_{\rm feat}} w_I\,
  \big\|\mathcal I_I(\sigma)-\mathcal I_I(\sigma')\big\|^2\right)^{1/2},\\
  \Ksig(\sigma,\sigma') &\equiv \exp\!\big[-d_\sigma(\sigma,\sigma')/\lambda_\sigma\big].
\end{split}
\tag{7.1}
\end{equation}

This makes $\Ksig(\sigma,\sigma')$ computable and controls nonlocality via $\lambda_\sigma$.

\textbf{Feature examples $\mathcal I_I$ (local, gauge/diffeo-invariant scalars):}
\begin{center}
\begin{tabular}{ll}
\toprule
Feature & Definition \\
\midrule
$\mathcal I_1$ & $F_{\mu\nu}F^{\mu\nu}$ (EM field strength squared) \\
$\mathcal I_2$ & $\bar\psi\gamma^\mu D_\mu\psi$ (fermion current density) \\
$\mathcal I_3$ & $R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}$ (curvature invariant) \\
$\mathcal I_4$ & $\nabla_\mu\phi\nabla^\mu\phi$ (scalar kinetic term) \\
\bottomrule
\end{tabular}
\end{center}

Each $\mathcal I_I$ is built from fields and a \textbf{finite number of derivatives}, ensuring locality in $M$. The number of features $N_{\rm feat}$ is finite for EFT control.

\textbf{Gauge and diffeomorphism invariance:} All features $\mathcal I_I$ and the selection operator $O_S$ are constructed from \textbf{gauge-invariant, diffeomorphism-covariant scalars}. For example, in the electromagnetic sector, one may use $\mathcal I_{\rm EM} = F_{\mu\nu}F^{\mu\nu}$ where $F_{\mu\nu}=\partial_\mu A_\nu - \partial_\nu A_\mu$ is the field strength tensor, which is manifestly gauge-invariant under $A_\mu \to A_\mu + \partial_\mu\chi$. Similarly, curvature scalars like $R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}$ are diffeomorphism-invariant. This ensures the framework respects the symmetries of general relativity and the Standard Model.

\subsection{Alternative: Signature-Based Extraction}

From local state dynamics $(s)$ (time series, spatial configuration):

\begin{itemize}
\item \textbf{Spectral:} FFT $\to$ power spectrum $\to$ dominant frequencies $(f_1, \ldots, f_m)$, spectral entropy $H_{spec}$,
\item \textbf{Topological:} persistent homology $\to$ Betti curves $(\beta_0(r), \beta_1(r))$,
\item \textbf{Statistical:} autocorrelation $\tau_c$, Lyapunov exponents $\lambda_i$ (if applicable), moments $(\mu_k)$.
\end{itemize}

Signature: $(\sigma(s) = (f_i, H_{spec}, \beta_j(r_k), \tau_c, \lambda_i, \mu_k) \in \mathbb{R}^d)$.

\subsection{Metric Definition}

\begin{equation}
d_\sigma(s,s') := \ell_0 \left[ \|\sigma(s) - \sigma(s')\|_2 + \alpha_W \, W(\mu_s, \mu_{s'}) \right],
\tag{7.1}
\end{equation}

where $(W)$ is the Gromov-Wasserstein distance between probability distributions $(\mu_s, \mu_{s'})$, $(\alpha_W \geq 0)$, and the \textbf{embedding scale $(\ell_0)$} gives $(d_\sigma)$ \textbf{length dimension} (meters). Typically $(\ell_0 \sim 1~\mu\text{m})$.

\subsection{Distance Ladder (Calibration)}

Construct controlled levels:

\begin{table}[h]
\centering
\begin{tabular}{clll}
\toprule
Level & Transformation & Expected $d_\sigma$ & Expected K \\
\midrule
0 & Identical $(s' = s)$ & 0 & 1 \\
1 & Phase rotation (spectrum preserved) & $\varepsilon_1 \ll \lambda_\sigma$ & $\approx 0.9$ \\
2 & Permuted label & $\varepsilon_2 \approx 0.3 \lambda_\sigma$ & $\approx 0.7$ \\
3 & Block-scramble (temporal/spatial) & $\varepsilon_3 \approx 0.7 \lambda_\sigma$ & $\approx 0.5$ \\
4 & Additive noise (SNR = 10 dB) & $\varepsilon_4 \approx \lambda_\sigma$ & $\approx 0.37$ \\
5 & Independent realization & $\varepsilon_5 \gg \lambda_\sigma$ & $\ll 0.1$ \\
\bottomrule
\end{tabular}
\end{table}

Test that $(\Ksig(e) = \exp[-d_\sigma/\lambda_\sigma])$ falls monotonically and that the effect in E1 correlates with level. This directly links observable effect to a calibrated scale.

\textbf{Pre-registration:} The distance ladder is pre-registered as a \textbf{separate, blind pilot} (own OSF DOI) before main experiments E1/E2/E3. Blind ``level labels'' are kept sealed until analysis phase.

