\section{Limitations, Open Questions, and Future Work}

\textbf{Limitations:}

\begin{itemize}
\item \textbf{$\alpha$-value:} We set $\alpha\equiv1$ for consistency with the Bianchi identity; this places the entire FTL mechanism in S-locality.
\item \textbf{S-mediator implementation:} Requires substrate with dynamic field $\chi(\sigma,T)$ $\to$ speculative.
\item \textbf{Q-factor in practice:} Unknown whether $(Q \sim 10^{-2})$ is achievable in realistic systems (BEC, Josephson, photonics near band-edge).
\item \textbf{Naturalness problem:} If $(\varepsilon \ll 1)$, why not $(\varepsilon = 0)$ exactly? New symmetry principle needed that permits but suppresses resonance.
\end{itemize}

\textbf{Open questions:}

\begin{enumerate}
\item \textbf{Substrate specification:} Which explicit rule set (cellular automaton, hypergraph rewriting, etc.) yields:
   \begin{itemize}
   \item Emergent Lorentz symmetry at low energy,
   \item Selection operator $(O_S)$ with $\Delta > 4$,
   \item Reduced gravitational coupling $(\alpha \ll 1)$?
   \end{itemize}

\item \textbf{Q-platforms:} Experimental mapping of $(Q)$ in:
   \begin{itemize}
   \item Josephson junctions near superradiance transition,
   \item Photonic crystals near band-edge,
   \item BEC near phase separation.
   \end{itemize}

\item \textbf{Entropy accounting:} Detailed model for how $(\Sigma_S)$ couples to physical heat bath; is MDL proxy sufficient or is algorithmic information theory (AIT) required?

\item \textbf{Coupling to emergent gravitation:} Can the $\alpha$-factor be derived from an induced-gravity mechanism where S-sector couples more weakly to metric emergence?
\end{enumerate}

\textbf{Future work:}

\begin{itemize}
\item \textbf{Explicit substrate model:} Construct a toy model (2D/3D cellular automaton) demonstrating emergent Lorentz + resonance.
\item \textbf{Material screening:} Identify candidate platforms with high-Q near criticality.
\item \textbf{First E1 test:} Pilot study with electrical RC networks (low cost) for proof-of-concept.
\end{itemize}

