\section*{Appendix E: Assumptions (Summary Checklist)}

\begin{itemize}
\item \textbf{(A1)} Global ordering $\tau$ with strict retardation (substrate time $T$ increases monotonically)
\item \textbf{(A2)} $\tilde{\mathcal{K}} \geq 0$ (resource monotonicity - no negative costs)
\item \textbf{(A3)} Sparse and weak S-links (max degree $g \ll N$, total strength $\eta$ small)
\item \textbf{(A4)} $O_S$ is RG-irrelevant ($\Delta > 4$) and $\langle O_S \rangle \approx 0$ in homogeneous states (degeneracy dilution)
\item \textbf{(A5)} $\mathbb{K}$ is positive semidefinite and causal in $\tau$ (ensures well-defined propagator)
\item \textbf{(A6)} $c_T = c$ in absence of resonance (gravitational waves travel at light speed)
\item \textbf{(A7)} Momentum-neutrality: $\int J^i_\sigma d^4x = 0$ (no reactionless drive)
\item \textbf{(A8)} Gravitational coupling: $\alpha = 1$ exactly (metric responds light-speed causally)
\item \textbf{(A9)} Length dimensions: $d_\sigma$ and $\lambda_\sigma$ have meters via embedding scale $\ell_0 \sim 1$ $\mu$m
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section*{Method Note (Experimental Best Practices)}

For proposals of this type (small signals, extraordinary claims):

\begin{enumerate}
\item \textbf{Pre-registration:} Open Science Framework or equivalent, SHA-256 hash of analysis code before data
\item \textbf{Blinding:} Seal critical parameters (distance ladder labels, sidereal phase) until analysis
\item \textbf{Environmental isolation:} Triple Faraday, optical isolation, battery power, vibration isolation
\item \textbf{Independent replication:} Ideally 3+ teams, different hemispheres, compare results
\item \textbf{Open data/code:} Publicly available after publication, allow independent analysis
\item \textbf{Statistical rigor:} Multiple-test correction (Holm-Bonferroni or FDR), permutation tests $\geq 10^6$
\item \textbf{Adversarial review:} Invite skeptics to design null tests, offer co-authorship
\end{enumerate}

All predictions must be quantitative. All null results must translate to parameter bounds with explicit error bars.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{This document describes a speculative but internally consistent mechanism with explicit falsifiability criteria. Either it results in stringent upper bounds - or in a new class of reproducible non-local effects. Both outcomes deserve careful testing.}

\textbf{Version:} popular1.md (Extended Popular Science Edition)\\
\textbf{Date:} 2025\\
\textbf{Status:} Theoretical proposal awaiting experimental test

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Further Reading:}

For technical details, see \textbf{article3.md} (the full mathematical formulation).

For experimental protocols, see \textbf{\S13} (this document) or contact the authors.

For philosophical implications, see:
\begin{itemize}
\item Wheeler, ``It from Bit'' (1990)
\item Tegmark, ``Mathematical Universe Hypothesis'' (2008)
\item Wolfram, ``A New Kind of Science'' (2002)
\end{itemize}

For related experimental work:
\begin{itemize}
\item Lieb-Robinson bounds: Hastings \& Koma (2006)
\item SME tests: Kosteleck\'y et al. (ongoing)
\item Emergent gravity: Verlinde (2011), Jacobson (1995)
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Acknowledgments:}

To the reader who made it this far: Thank you for engaging with these ideas. Whether you're convinced, skeptical, or just curious - your critical thinking is what science needs.

Now let's go build some experiments and find out if reality has a basement.

