\section{Limitations and Open Questions}

\begin{quote}
\textbf{Core Concept:} Science is about honesty. Here's what we don't know and what might be wrong.
\end{quote}

\subsection{Known Limitations}

\textbf{1. $\alpha$-value constraint:}\\
We set $\alpha \equiv 1$ for consistency with Bianchi identity. This is derived, not assumed, but it places the entire FTL mechanism in S-locality. If $\alpha \neq 1$, the framework needs modification.

\textbf{2. S-mediator implementation:}\\
The $\chi(\sigma,T)$ field is speculative. We don't have an explicit substrate model showing how it emerges. This is like proposing atoms before knowing about quarks and electrons - the framework is consistent, but details are missing.

\textbf{3. Q-factor in practice:}\\
We don't know if $\mathcal{Q} \sim 10^{-2}$ is achievable. Current quantum systems (BECs, superconductors, photonic cavities) have quality factors, but not necessarily the right kind for aether resonance.

\textbf{Best candidates:}
\begin{itemize}
\item Josephson junctions near superradiance transition
\item Photonic crystals near band-edge (slow light)
\item BEC near phase separation
\item Cold atoms in optical lattices near quantum critical point
\end{itemize}

All speculative - needs experimental investigation.

\textbf{4. Naturalness problem:}\\
If $\varepsilon \ll 1$ ($\sim 10^{-15}$), why isn't $\varepsilon = 0$ exactly?

\emph{Analogy:} Why is the electron mass 511 keV and not 0 or 1000 GeV? Naturalness says: ``If a parameter is small, there should be a symmetry explaining why.''

\textbf{Possible answer:} A new symmetry that's approximate (like chiral symmetry for quarks). The symmetry forbids resonance exactly, but it's spontaneously broken at some scale, generating $\varepsilon \neq 0$ but small.

This is hand-waving - we need a concrete model.

\subsection{Open Questions (Future Work)}

\textbf{1. Explicit substrate specification:}\\
Which rule-set (cellular automaton, hypergraph rewriting, spin network evolution) gives:
\begin{itemize}
\item Emergent Lorentz symmetry at low energy
\item Selection operator $O_S$ with $\Delta > 4$
\item Weak resonance coupling $\varepsilon \sim 10^{-15}$
\end{itemize}

This is the ``Holy Grail'' - an explicit toy model we can simulate.

\textbf{2. Q-platform mapping:}\\
Systematic experimental survey:
\begin{itemize}
\item Measure $\mathcal{Q}$ in various quantum platforms
\item Look for signatures of structural proximity effects
\item Map phase diagram: $\mathcal{Q}$(temperature, drive, detuning)
\end{itemize}

\textbf{3. Entropy bookkeeping:}\\
Detailed model for how $\Sigma_S$ couples to physical heat bath:
\begin{itemize}
\item Is MDL proxy sufficient?
\item Or do we need full algorithmic information theory (AIT)?
\item Can we derive (8.2) from microscopic master equations?
\end{itemize}

\textbf{4. Coupling to emergent gravitation:}\\
Can $\alpha = 1$ be derived from induced-gravity mechanisms?
\begin{itemize}
\item Idea: Spacetime curvature emerges from substrate entanglement (like Jacobson, Verlinde)
\item S-sector entanglement couples more weakly
\item Result: $\alpha \sim$ (S-coupling)/(M-coupling) set by entanglement structure
\end{itemize}

This connects to quantum gravity programs (emergent gravity, entropic gravity).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

