\documentclass[11pt,a4paper]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{physics}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}

\geometry{margin=1in}

% === Canonical macros shared across article3.tex and popular1.tex ===
% Spaces & projections
\newcommand{\M}{\mathcal{M}}          % emergent spacetime
\newcommand{\Ssub}{\mathcal{S}}        % pattern/substrate space
\newcommand{\proj}{\pi}                % projection \Ssub -> \M
% Distances & scales
\newcommand{\ds}{d_{\sigma}}
\newcommand{\dsig}{d_{\sigma}}
\newcommand{\ls}{\lambda_{\sigma}}
\newcommand{\lsig}{\lambda_{\sigma}}
\newcommand{\lzero}{\ell_{0}}
% Kernels & operators
\newcommand{\Kern}{\mathcal{K}}        % bilocal resonance kernel on \Ssub
\newcommand{\Ksig}{K}    % similarity factor K = exp(-d_sigma/lambda_sigma)
\newcommand{\Osel}{\mathcal{O}_{\!S}}  % selection operator (dim 4)
\newcommand{\Os}{O_{S}}                % alternate form
% Substrate time & temperature
\newcommand{\Tsub}{T}                  % substrate ticks
\newcommand{\Temp}{\mathcal{T}}        % temperature
% Substrate current and scalar power
\newcommand{\Jsig}{J^{\nu}_{\sigma}}   % 4-current in (3.4)-(3.6)
\newcommand{\Psig}{P_{\sigma}}         % scalar power flow used in Sec. 6
% Æther / clock
\newcommand{\taufield}{\tau}
\newcommand{\uvec}{u^{\mu}}
\newcommand{\lamL}{\lambda_{L}}        % Lagrange multiplier (replaces \Lambda(x))
\newcommand{\LamUV}{\Lambda_{\!*}}     % UV scale in Sint
% Gravitational coupling
\newcommand{\alphaG}{\alpha_{\!g}}     % gravitational coupling (not fine-structure!)
% Efficiencies & sector labels
\newcommand{\betaBER}{\beta_{\mathrm{BER}}}
\newcommand{\Qmat}{Q_{\mathrm{mat}}}
\newcommand{\Qgam}{Q_{\gamma}}
\newcommand{\epsmat}{\varepsilon_{\mathrm{mat}}}
\newcommand{\epsgam}{\varepsilon_{\gamma}}
\newcommand{\PumpRate}{\Gamma_{\mathrm{pump}}} % = P_pump/(\hbar\omega_0)
% Legacy/compatibility
\newcommand{\Kkernel}{\mathbb{K}}
\newcommand{\Krate}{\tilde{\mathcal{K}}}
\newcommand{\fl}{f_{\ell}}
\newcommand{\veff}{v_{\!eff}}
\newcommand{\Ts}{T^{\mu\nu}_{S}}
\newcommand{\Asid}{A_{\rm sid}}

% === Symbol tables and assumptions are in the appendices ===

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\title{Aether Resonance: How Reality Might Cheat at Its Own Speed Limit\\[0.5em]
\large A Guide for the Curious Mind}
\author{}
\date{}

\begin{document}

\maketitle

\noindent\textbf{What you're about to read:} This is a scientific proposal for how the universe might allow faster-than-light communication without breaking the laws of physics. We keep all the mathematical details, but we'll walk through them together, one step at a time, using everyday analogies and concrete examples. Think of this as a guided tour through a radical idea about how reality works at its deepest level.

\begin{center}
\rule{0.5\linewidth}{0.5pt}
\end{center}

\section*{Abstract (The Elevator Pitch)}

Imagine you're playing a massive multiplayer video game. The game has rules - you can't walk through walls, gravity pulls you down, light travels at a certain speed through the game world. These are the ``emergent'' rules you experience as a player.

But underneath, there's a computer running the game. It has its own structure - CPUs, memory, data structures - that operates by completely different rules. Two characters separated by a huge in-game distance might have their data stored right next to each other in the computer's memory.

We're proposing that our universe might work something like this. The ``game rules'' are Einstein's relativity and quantum mechanics - they work perfectly and give light its cosmic speed limit. But perhaps there's a computational ``substrate'' underneath, with its own notion of what ``nearby'' means. If two things are similar enough in structure (like two identical quantum systems), they might be ``close'' in this deeper sense, even if they're far apart in ordinary space.

This hypothesis proposes \textbf{aether resonance} - a mechanism for faster-than-light communication that exploits this substrate-level closeness while respecting all the regular laws of physics in the emergent spacetime we inhabit.

\textbf{What we provide:}
\begin{enumerate}
\item A complete mathematical framework showing how this could work
\item Explanations of why it doesn't create paradoxes or violate conservation laws
\item Three specific experiments you could actually build to test it
\item Predictions of what would happen if the idea is right (or wrong)
\end{enumerate}

\begin{center}
\rule{0.5\linewidth}{0.5pt}
\end{center}

\input{popular1-sec01.tex}
\input{popular1-sec02.tex}
\input{popular1-sec03.tex}
\input{popular1-sec04.tex}
\input{popular1-sec05.tex}
\input{popular1-sec06.tex}
\input{popular1-sec07.tex}
\input{popular1-sec08.tex}
\input{popular1-sec09.tex}
\input{popular1-sec10.tex}
\input{popular1-sec11.tex}
\input{popular1-sec12.tex}
\input{popular1-sec13.tex}
\input{popular1-sec14.tex}
\input{popular1-sec15.tex}
\input{popular1-sec16.tex}

\input{popular1-appA.tex}
\input{popular1-appB.tex}
\input{popular1-appC.tex}
\input{popular1-appD.tex}
\input{popular1-appE.tex}
\input{popular1-appF.tex}

\end{document}
